Cluster Difference Imaging Photometric Survey. III. An Open Cluster Spread Across 500\,pc

### 2021-03-12 cluster_paper/ms.tex
update ms

Gaia-ESO selected candidate \cn\ members to observe with the GIRAFFE
and UVES spectrographs based on previously reported literature members
and publicly available photometry.  Since the existence of the \cn\
halo was not known at the time of target selection, very few ``halo''
stars are in the sample.  For \cn, \citet{randich_gaiaeso_2018}
ultimately reported stellar parameters (including lithium equivalent
widths and metallicities) for 796 stars that they considered possible
\cn\ members.  Cross-matching against our kinematic list of
\nkinematic candidate members by position and imposing a 0$\farcs$5
maximum separation limit yields 492 kinematic members for which a
Gaia-ESO spectrum was acquired and stellar parameters were reported.
15 of these (all with $(Bp-Rp)_0 > 2.0$) are spurious matches based on
the Gaia color and GES effective temperature, and we remove them.
This yields 477 stars, of which 436 are in the core, and 41 are in the
halo.  The lop-sided ratio is due to the Gaia-ESO selection function.
We verified by directly querying the ESO archive that
\citet{randich_gaiaeso_2018} indeed included $>99\%$ of the available
spectra as part of their \cn\ sample, and so we opted to use their
results rather than repeating their analysis.

% GALAH DR3 selected XXX/YYY.  GALAH-DR3 = GALAH-main + GALAH-faint +
% K2-HERMES + TESS-HERMES + HERMES other.  Data taken both with the 2dF
% fiber positioner, and the HERMES spectrograph at the 3.9 AAT
% (Lewis+02, Sheinis+15).  GALAH-main is 12<V<14, $\delta<+10^\circ$,
% $|b|>10^\circ$, in regions of sky with at least 400 targets in $\pi$
% square degrees (the 2dF FoV).  Fields with more than 400 targets were
% targetted multiple times, with different fiber arrangements.
% TESS-HERMES: 10.0 < V < 13.1. TESS Southern CVZ, within $12^\circ$ of
% the southern ecliptic pole.  54354 in the ``HERMES other'' program are
% from targeted observations of stars in open clusters, the GALAH Pilot
% Survey (Martell et al. 2017),

@@

%
%/Users/luke/Dropbox/proj/earhart/results/lithium/kinematic_X_galah_dr3.csv
%
% https://docs.datacentral.org.au/galah/dr3/catalogue-data-access/
We downloaded\footnote{Via \url{datacentral.org.au/services/download},
using the \texttt{sobject\_id} identifiers.} the GALAH DR3 spectra for
all 107 entries.  We then measured the lithium equivalent widths using
the AstroPy-affiliated \texttt{specutils} package
\citep{specutils_v1pt1}.

%FIXME TODO: put it into words
% circa ~line 1224 from ~/Dropbox/proj/cdips_followup/cdips_followup/spectools.py
Our procedure was as follow...

We then measured the lithium equivalent widths using PROCESS Y.
We then concatenated the GALAH DR3 and Gaia-ESO results together.

% TODO: fix the fact that the figure currently only shows Randich+18
Figure~\ref{fig:lithium} shows the results.  At fixed stellar mass
(and age), the rapid rotators tend to show elevated lithium equivalent
widths. This effect is mostly apparent in the K dwarfs.  Similar
trends in the Pleiades were noted nearly three decades ago by
\citet{soderblom_evolution_1993}.  More recent studies have found
comparable results in the Pleiades, the Psc-Eri stream, and M\,35
\citep{bouvier_pleiades_lirot_2018,arancibia_2020,jeffries_m35_li_2020}.
The fact that we see the same trend across the core and halo of \cn\ 
{\it a)} supports the conclusion that the halo is coeval with the
core, and {\it b)} suggests that the lithium-rotation correlation is
not explained by environmental effects such as the density of the cluster
during its embedded phase, but is instead tied to physical processes
acting on all low-mass pre-main-sequence stars.

@@

The authors thank X and Y for fruitful discussions.
%
L.G.B. and J.H. acknowledge support by the TESS GI Program, program
NUMBER, through NASA grant NUMBER.
L.G.B. was also supported by a XXXX Fellowship from Princeton
University.
%
This study was based in part on observations at Cerro Tololo
Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop. ID
2020A-0146; 2020B-NUMBER PI: L{.}~Bouma), which is managed by the
Association of Universities for Research in Astronomy (AURA) under a
cooperative agreement with the National Science Foundation.
%
ACKNOWLEDGE PFS / CAMPANAS.
%
This paper includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
We thank the TESS Architects (George Ricker, Roland Vanderspek, Dave
Latham, Sara Seager, Josh Winn, Jon Jenkins) and the many TESS team
members for their efforts to make the mission a continued success.
%

>>>>

Gaia-ESO selected candidate \cn\ members to observe with the GIRAFFE
and UVES spectrographs based on previously reported literature members
and publicly available photometry.  Since the existence of the \cn\
halo was not known at the time of target selection, very few ``halo''
stars are in the sample.  For \cn, \citet{randich_gaiaeso_2018}
ultimately reported stellar parameters (including lithium equivalent
widths and metallicities) for 796 stars that they considered possible
\cn\ members.  Cross-matching against our kinematic list of
\nkinematic candidate members by position and imposing a 0$\farcs$5
maximum separation limit yields 492 kinematic members for which a
Gaia-ESO spectrum was acquired and stellar parameters were reported.
15 of these (all with $(Bp-Rp)_0 > 2.0$) are spurious matches based on
the Gaia color and GES effective temperature, and we remove them.
This yields 477 stars, of which 436 are in the core, and 41 are in the
halo.  The lop-sided ratio is due to the Gaia-ESO selection function.
We verified by directly querying the ESO archive that
\citet{randich_gaiaeso_2018} indeed included $>99\%$ of the available
Gaia-ESO spectra as part of their \cn\ sample, and so we opted to use
their results rather than repeating their analysis.

@@

%
%/Users/luke/Dropbox/proj/earhart/results/lithium/kinematic_X_galah_dr3.csv
%
% https://docs.datacentral.org.au/galah/dr3/catalogue-data-access/
We downloaded\footnote{Via \url{datacentral.org.au/services/download},
using the \texttt{sobject\_id} identifiers.} the GALAH DR3 spectra for
all 107 entries.  We then measured the lithium equivalent widths using
the \texttt{specutils} package \citep{specutils_v1pt1}.  We did this
by first focusing in on a 15\AA\ window centered on the 6707.835\AA
lithium doublet.  We then continuum normalized the spectra using a
third-degree Chebyshev series, while excluding any regions that showed
absorption lines ({\it e.g.}, Fe\,\textsc{I} is present at
6703.58\,\AA and 6705.10\,\AA).  We proceeding by fitting a Gaussian
to the continuum normalized spectra, considering only a $\pm 1\,$\AA
window centered on the Li doublet.  The equivalent widths were then
evaluated by numerically integrating the fitted model over the same
window.  Our approach therefore includes the Fe\,6707.44 blend in the
reported Li equivalent widths---which leads to systematic
overestimates of $\approx$10 to 15\,m\AA\ \citep[{\it
e.g.},][]{bouvier_pleiades_lirot_2018}.  Finally, to derive
uncertainties on the EWs, we repeated the procedure twenty times,
but added noise to the spectra, drawn from a normal distribution with
a scale set by the standard deviation of the continuum.
The reported uncertainties are then drawn from the $84^{\rm th}$ and
$16^{\rm th}$ percentiles of the resulting EW distribution.
We verified the overall scale of our results by comparing the 21
stars that overlapped between GALAH and Gaia-ESO.
17 of the 21 measurements agreed within 1-$\sigma$.
The remaining four were stars with $T_{\rm eff}>6200\,{\rm K}$, for
which \citet{randich_gaiaeso_2018} reported Li EW detections
ranging from $50$ to $75\,$m\AA, and for which we found 
either null results or only slightly significant
($\approx$30m\AA) detections.

% The outliers:
%    Fitted_Li_EW_mA     EWLi            source_id  Teff
%    24           -1.490   76.100  5290730488348194304  6786
%    34           19.485   71.300  5290935341107132800  6702
%    15           20.251   50.867  5290722310730727040  6197
%    25           29.378   52.500  5290724681552584192  6511

Figure~\ref{fig:lithium} shows the results.  
At fixed stellar mass
(and age), the rapid rotators tend to show elevated lithium equivalent
widths. This effect is mostly apparent in the K dwarfs.  Similar
trends in the Pleiades were noted nearly three decades ago by
\citet{soderblom_evolution_1993}.  More recent studies have found
comparable results in the Pleiades, the Psc-Eri stream, and M\,35
\citep{bouvier_pleiades_lirot_2018,arancibia_2020,jeffries_m35_li_2020}.
The fact that we see the same trend across the core and halo of \cn\ 
{\it a)} supports the conclusion that the halo is coeval with the
core, and {\it b)} suggests that the lithium-rotation correlation is
not explained by environmental effects such as the density of the cluster
during its embedded phase, but is instead tied to physical processes
acting on all low-mass pre-main-sequence stars.

@@

The authors thank X and Y for fruitful discussions.

B.~Tofflemire, % lithium suggestoins

%
L.G.B. and J.H. acknowledge support by the TESS GI Program, program
NUMBER, through NASA grant NUMBER.
L.G.B. was also supported by a XXXX Fellowship from Princeton
University.
%
This study was based in part on observations at Cerro Tololo
Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop. ID
2020A-0146; 2020B-NUMBER PI: L{.}~Bouma), which is managed by the
Association of Universities for Research in Astronomy (AURA) under a
cooperative agreement with the National Science Foundation.
%
ACKNOWLEDGE PFS / CAMPANAS.
%
This paper includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
We thank the TESS Architects (George Ricker, Roland Vanderspek, Dave
Latham, Sara Seager, Josh Winn, Jon Jenkins) and the many TESS team
members for their efforts to make the mission a continued success.
%
